Fail Faster

### 2025-03-15 paper/paper.tex
blahg

\jwc{Unclear what we should call this section}

As we discussed in Section~\ref{section:motiv}, inefficient RNG choice is
another bottleneck for finding bugs fast. While generator libraries by
and large use sensible RNG algorithms, they are not explicitly chosen
with performance in mind. Indeed, much faster RNGs \cn and fast sources
of entropy \cn exist. \jwc{... more here?}
To demonstrate that a faster RNG can significanty impact bugfinding power, we
use the natural experiment provided by OCaml's inefficient implementation of
SplitMix. By replacing this slow RNG with a faster but extensionally equivalent
implemetnation, we are able to precisely quantify the bugfinding speedup that using a RNG gives across a range of PBT scenarios.

@@

The precise details of how the SplitMix algorithm works are unimportant for the
present paper, but the critical component is that all of its operations are are
defined in terms of arithmetic bitwise operations on 64-bit integers.  In OCaml,
a
% however,

% this has a significant impact on the performance of generators. a single call to
% one of the \bq library functions may sample from \texttt{splittable\_random}
% multiple times, and each sample from \texttt{splittable\_random} allocates 9
% words \jwc{this is a call to \texttt{next\_int64}}.  while seemingly small, we
% will see in section~\ref{section:eval} that this amount of allocation has a
% significant impact on significant.

\jwc{
  \begin{itemize}
    \item as we discussed in section~\ref{section:motiv}, 
    \item faster rngs and sources of entropy exist, though they are extensionally inequivalent. 
    \item 
    \item 
    \item here's how splitmix is slow: the core of splitmix is nextint64, which does allocations for each operation.
    \item if we could use actual int64s, we wouldn't have to allocate.
    \item so we do this:
    \begin{itemize}
      \item we implement splitmix in c (which uses normal int64s) and real int64 arithmetic, and use the ocaml ffi to call out to the c functions. t
      \item this could be done with jane street bleeding-edge compiler extensions, though these are not yet implemented in a a version of the ocaml compiler that also has staging.
    \end{itemize}
  \end{itemize}
}

>>>>

\jwc{Unclear what we should call this section}
As we discussed in Section~\ref{section:motiv}, inefficient RNG choice is
another bottleneck for finding bugs fast. While generator libraries by
and large use sensible RNG algorithms, they are not explicitly chosen
with performance in mind. Indeed, much faster RNGs \cn and fast sources
of entropy \cn exist. \jwc{... more here?}
To demonstrate that a faster RNG can significanty impact bugfinding power, we
use the natural experiment provided by OCaml's inefficient implementation of
SplitMix. By replacing this slow RNG with a faster but extensionally equivalent
implemetnation, we are able to precisely quantify the bugfinding speedup that using a RNG gives across a range of PBT scenarios.

@@

The precise details of how the SplitMix algorithm works \cn are unimportant for the
present paper, but the critical component is that all of its operations are are
defined in terms of arithmetic bitwise operations on 64-bit integers. In OCaml,
because of details related to the garbage collector, the 64-bit integer type \texttt{int64} is represented
at run time as a \emph{pointer} to an unscanned block of memory containing (among other things)
a 64-bit integer \cn.

This means that all operations that return an \texttt{int64} must allocate this block of memory.
This has a significant impact on the performance of generators. A single call to
one of the \bq library functions --- like generating an integer uniformly in a range --- may call the \texttt{splittable\_random}
algorithm multiple times. Each sample from \texttt{splittable\_random} allocates 9
words \jwc{this is a call to \texttt{next\_int64}}, and each allocation brings us closer
to the next garbage collection pause. While allocation in OCaml is fast and the garbage colletor
runs incrementally in small slices
\footnote{The OCaml GC is a generational collector \cn, and since these allocations are mostly very short lived, they will never be promoted. }, we will see in Section~\ref{section:eval} that this can have
a large performance impact on some generators that spend most of their time sampling data.

% \jwc{
%   \begin{itemize}
%     \item as we discussed in section~\ref{section:motiv}, 
%     \item faster rngs and sources of entropy exist, though they are extensionally inequivalent. 
%     \item here's how splitmix is slow: the core of splitmix is nextint64, which does allocations for each operation.
%     \item if we could use actual int64s, we wouldn't have to allocate.
%     \item so we do this:
%     \begin{itemize}
%       \item we implement splitmix in c (which uses normal int64s) and real int64 arithmetic, and use the ocaml ffi to call out to the c functions. t
%       \item this could be done with jane street bleeding-edge compiler extensions, though these are not yet implemented in a a version of the ocaml compiler that also has staging.
%     \end{itemize}
%   \end{itemize}
% }
